Numerical error propagation in the HCP structural pre-processing pipelines

### 2018-03-15 paper.tex
Wrote in section 2.5

@@

\subsection{Pipeline analysis: Characterizing differences}

Once differences between conditions are quantified, we automatically
identify the steps in the pipeline responsible for such differences.
We use the \reprozip tool~\cite{reprozip} to record: (1) the tree of
processes executed by the pipeline and (2) the list of files read and
written by each process. This information is collected by system call
interceptions, through the \texttt{ptrace} as a UNIX base system call, and stored in a \texttt{SQLite}
database. The database stores information on all the processes which
are created by the \texttt{clone()} in Linux or \texttt{fork()} in UNIX-like
operating system calls. It also records the files opened by each process (in read, write or execution mode), including the final and
temporary files.

We reconstruct the tree of processes starting from the first process
created by the pipeline and defining its child processes as the ones
that were created through \texttt{clone()} or \texttt{fork()}. We
create a process graph from this tree by adding edges corresponding to
file dependencies between processes. A file dependency is defined
between processes A and B if a file written by A is read by
B. Figure~\ref{fig:graph-example} shows an example of a process graph
constructed at this stage.

\begin{figure}
%  \includegraphics{brain\_classification}
  \includegraphics[scale=0.3]{images/simple_graph}
  \caption{Process graph constructed from an example brain extraction script.}
  \label{fig:graph-example}
\end{figure}

A different process graph may be produced for every subject in every
execution condition. Among our dataset, we identified 4 types of
subjects with different numbers of T1w and T2w images. We verified
that the process graphs were identical for all subjects of the same
type, for all execution conditions. The remainder of the analysis is
done separately for each subject type.

Cycles may be present in the process graph in case a file was written
by more than one process.  We remove the cycles in the process graph
by removing file edges between processes A and B when A's process
creation timestamp is posterior to B's or when A=B.

\todo{Start from here}
\paragraph{} In the next step, we classify the processes into three categories considering to the obtained process tree 
and the error matrix file. First, processes that read file(s) without error but write different file(s) create error in 
the pipeline. The second group of the processes are those which remove errors in pipeline, read file(s) with error and
 write “nothing or” file(s) without error. Most of the processes are in the third category beyond any fault which
 read and write file(s) without error. The processes in each category may read or write temporary file(s) that implies 
the uncertainty of the process.
\paragraph{} Pipeline is conducted of a bunch of processes sequentially, setting a modification step after classification
 of processes iteratively enables us to control the propagation of errors in the pipeline correctly. In each iteration
 we detect all the processes which are create error in the pipeline and then by fixing those processes artificially, 
we’ll find the other processes in the pipeline that introduce an error. This procedure will be terminated by fixing all the processes. 
To fix the processes that create error, we replace their dependency files with error by the same file copy from the results 
obtained in the other operating system. In this way, we were able to pinpoint the origin of the errors measured in the pipeline.

\todo{Check that} We also recorded the provenance trace of the execution using \reprozip. \reprozip identifies all the
processes and associated parameters that wrote the files having
differences.  These details about various processes is helpful in
debugging the pipelines. This information helps in recreating the
processing step by step and also to identify the processes that
creates the differences.

>>>>

\todo{Check everywhere that differences are called error. And define that in the intro.}

@@

\subsection{Pipeline analysis: Characterizing errors}

Once errors between conditions are quantified, we automatically
identify the steps in the pipeline responsible for such differences.
We use the \reprozip tool~\cite{reprozip} to record: (1) the tree of
processes executed by the pipeline and (2) the list of files read and
written by each process. This information is collected by system call
interceptions, through the \texttt{ptrace} Unix system call, and
stored in a \texttt{SQLite} database. The database stores information
on all the processes which are created by the \texttt{clone()} or
\texttt{fork()} system calls. It also records the files opened by each
process (in read, write or execution mode), including the result files
and the temporary files.

We reconstruct the tree of processes starting from the first process
created by the pipeline and identifying its child processes as the ones
that were created through \texttt{clone()} or \texttt{fork()}. We
create a process graph from this tree by adding edges corresponding to
file dependencies between processes. A file dependency is defined
between processes A and B if a file written by A is read by
B. Figure~\ref{fig:graph-example} shows an example of a process graph
constructed at this stage.

\begin{figure}
%  \includegraphics{brain\_classification}
  \includegraphics[scale=0.3]{images/simple_graph}
  \caption{Process graph constructed from an example brain extraction
    script. Every node in the graph is labeled using (1) a process id
    created by our reconstruction, (2) the name of the executable run
    by the process. Processes that read or write temporary files are
    represented with squares; other ones with circles. Plain edges
    represent the process tree (\texttt{fork()} or \texttt{clone()}
    system calls). Dashed edges represent file dependencies: temporary
    files are in yellow and result files are in green.\todo{Add the
      script next to the graph. Remove node colors.}}
  \label{fig:graph-example}
\end{figure}

A different process graph may be produced for every subject in every
execution condition. Among our dataset, we identified 4 types of
subjects with different numbers of T1w and T2w images. We verified
that the process graphs were identical for all subjects of the same
type, for all execution conditions \todo{Check that}. The remainder of
the analysis is done separately for each subject type.

Cycles may be present in the process graph in case a file was written
by more than one process.  We remove the cycles in the process graph
by removing file edges between processes A and B when A's process
creation timestamp is posterior to B's or when A=B \todo{to be done}.

\paragraph{} In the next step, we classify the processes into three categories based on the process graph
and the error matrix file mentioned in the previous Section:
\begin{enumerate}
\item Processes that read files that do not have errors and write files that do not have errors are \emph{transparent}.
\item Processes that read files that do not have any error but write files that have errors \emph{create} errors in the pipeline. 
\item Processes that read files that have errors and write files that do not have errors \emph{remove} errors from the pipeline.
\end{enumerate}

The classification of processes that read or write temporary files is
uncertain when these files are deleted during the execution. In such
situations, our classification marks processes as uncertain as we were
unable to find a technical solution to prevent temporary files from
being deleted by the pipeline \todo{investigate that a bit more}.

Another challenge is to classify processes that read files that have
errors and write files that also have errors. In such situations, it
is not possible to determine from the pipeline results whether
the process created errors, removed some errors, or was
transparent. We label such processes \emph{unknown}. To address this issue, we developed an iterative approach that consists of the following steps:
\begin{enumerate}
  \item Run the pipeline in conditions 1 and 2; classify the
    processes as \emph{transparent}, \emph{create errors},
    \emph{remove errors} or \emph{unknown}.
  \item If there are \emph{unknown} processes then: replace each process P that create errors by a process Q that
    copies the results produced by process P in condition 1 to the pipeline output in condition 2.
  \item Repeat steps 1 and 2 until there is no \emph{unknown} process left.
\end{enumerate}
The replacement of process P at step 2 is done by overwriting P's
executable with a custom script produced by step 1. This custom script
copies the results obtained in condition 1 if it is invoked with the
arguments of a process that created errors, and calls P's original
executable otherwise. This algorithm converges to a process graph
without any \emph{unknown} process after a finite number of iterations.

Figure~\ref{fig:iterations} illustrates this iterative process on the previous
exemplar pipeline.
